\documentclass[openany]{book}
\input{../Main/Preamble.tex}
\begin{document}

\section{January 21, 2026}


\noindent\textbf{Plan:}
\begin{itemize}
  \item Translate section 2 of Svidzinsky 
  \item Establish a firm ground of accepted facts from which the BdG equations will follow.
  \item Understand how the approximation of the interaction is made.
\end{itemize}
\noindent\textbf{Evaluation:} 
I did find a way to translate section 2 of Svidzinsky, but the "accepted facts list" I generated needs improvement, and I did not get to bullet 3.

\section{January 22, 2026}
\noindent\textbf{Plan:} 
\begin{itemize}
  \item Do a better job identifying and understanding the necessary foundational material needed to understand the BdG derivation
\end{itemize}
\subsection{Fock space and Field operators}

\textbf{Physical meaning.}  
\(\mathcal H_1^{\wedge N}\) is the Hilbert space describing \(N\) identical fermions built from a one-particle Hilbert space \(\mathcal H_1\). A vector in \(\mathcal H_1^{\wedge N}\) encodes all probability amplitudes for finding the \(N\) fermions in arbitrary one-particle states, with the fermionic exchange sign built in. Fermionic Fock space collects all such \(N\)-particle sectors into a single Hilbert space.

\textbf{Mathematical definitions.}

\begin{enumerate}
\item \textbf{Permutation operators.}  
Let \(\mathcal H_1\) be a complex separable Hilbert space and let \(\mathcal H_1^{\otimes N}\) be its \(N\)-fold tensor product.  
For each permutation \(\pi \in S_N\), define a linear operator
\[
P_\pi : \mathcal H_1^{\otimes N} \to \mathcal H_1^{\otimes N}
\]
by its action on simple tensors,
\[
P_\pi(\phi_1 \otimes \cdots \otimes \phi_N)
= \phi_{\pi^{-1}(1)} \otimes \cdots \otimes \phi_{\pi^{-1}(N)},
\]
extended by linearity and continuity.

\item \textbf{Antisymmetrizer and its range.}  
Define the antisymmetrization operator
\[
A_N \equiv \frac{1}{N!}\sum_{\pi \in S_N} \mathrm{sgn}(\pi)\, P_\pi .
\]
The range of \(A_N\), denoted \(\mathrm{Ran}(A_N)\), is
\[
\mathrm{Ran}(A_N)
\equiv
\{\, A_N \Phi \mid \Phi \in \mathcal H_1^{\otimes N} \,\}.
\]
Since \(A_N^2 = A_N\) and \(A_N^\dagger = A_N\), \(\mathrm{Ran}(A_N)\) is a closed subspace of \(\mathcal H_1^{\otimes N}\).

\item \textbf{Fermionic \(N\)-particle Hilbert space.}  
The fermionic \(N\)-particle Hilbert space is defined as
\[
\mathcal H_1^{\wedge N} \equiv \mathrm{Ran}(A_N).
\]
Equivalently,
\[
\mathcal H_1^{\wedge N}
=
\left\{
\Psi \in \mathcal H_1^{\otimes N}
\;\middle|\;
P_\pi \Psi = \mathrm{sgn}(\pi)\,\Psi
\ \text{for all } \pi \in S_N
\right\}.
\]

\item \textbf{Combining one-particle states.}  
Given \(\phi_1,\ldots,\phi_N \in \mathcal H_1\), the corresponding \\ fermionic \(N\)-particle state is
\[
\phi_1 \wedge \cdots \wedge \phi_N
\equiv
A_N(\phi_1 \otimes \cdots \otimes \phi_N).
\]
This vector lies in \(\mathcal H_1^{\wedge N}\) and changes sign under exchange of any two factors.

\item \textbf{Meaning of an \(N\)-particle state.}  
For any \(\Psi \in \mathcal H_1^{\wedge N}\) and any \(\chi_1,\ldots,\chi_N \in \mathcal H_1\), the complex number
\[
\langle \chi_1 \wedge \cdots \wedge \chi_N , \Psi \rangle
\]
is interpreted as the probability amplitude that the \(N\) fermions occupy the one-particle states \(\chi_1,\ldots,\chi_N\).

\item \textbf{Algebraic sum and direct sum.}  
Consider the algebraic sum
\[
\sum_{N=0}^\infty \mathcal H_1^{\wedge N}
=
\left\{
\sum_{k=1}^m \Psi^{(N_k)}
\;\middle|\;
\Psi^{(N_k)} \in \mathcal H_1^{\wedge N_k},\ m<\infty
\right\}.
\]
This sum is direct because
\[
\mathcal H_1^{\wedge N} \cap \mathcal H_1^{\wedge M} = \{0\}
\quad \text{for } N \neq M .
\]

\item \textbf{Fermionic Fock space.}  
The fermionic Fock space is the Hilbert-space completion of this direct sum:
\[
\mathcal F \equiv \bigoplus_{N=0}^\infty \mathcal H_1^{\wedge N},
\]
with inner product
\[
\langle \Psi, \Phi \rangle
=
\sum_{N=0}^\infty
\langle \Psi^{(N)}, \Phi^{(N)} \rangle_{\mathcal H_1^{\wedge N}},
\quad
\Psi = \oplus_N \Psi^{(N)},\ 
\Phi = \oplus_N \Phi^{(N)}.
\]
\end{enumerate}

\subsection*{F2. Field operators on fermionic Fock space}


\subsubsection*{F2-1 Configuration-space representation, probability interpretation, and antisymmetry}

An element \(\Psi^{(N)}\in\mathcal H_1^{\wedge N}\) admits a coordinate–spin representation
\[
\Psi^{(N)}(x_1,\dots,x_N),
\qquad x_i=(\mathbf r_i,\sigma_i)\in\mathbb R^3\times\{\uparrow,\downarrow\},
\]
which is a complex-valued function on the \(N\)-particle configuration space
\((\mathbb R^3\times\{\uparrow,\downarrow\})^N\).

For any measurable region \(\Omega\) of this configuration space,
\[
\int_\Omega |\Psi^{(N)}(x_1,\dots,x_N)|^2\,dx_1\cdots dx_N
\]
is the probability of finding the \(N\) identical fermions with coordinates and spins in \(\Omega\).

Because the fermions are physically identical, \(\Psi^{(N)}\) satisfies the antisymmetry condition
\[
\Psi^{(N)}(x_{\pi(1)},\dots,x_{\pi(N)})
=
\mathrm{sgn}(\pi)\,\Psi^{(N)}(x_1,\dots,x_N)
\quad\text{for all }\pi\in S_N,
\]
which is the coordinate-space representation of the abstract relation
\(P_\pi\Psi^{(N)}=\mathrm{sgn}(\pi)\Psi^{(N)}\).
This antisymmetry leaves probabilities invariant while encoding fermionic exchange at the level of amplitudes.

\subsubsection*{F2-2 Definition of annihilation and creation operators and induced field operators}

For each \(f\in L^2(\mathbb R^3)\) and spin \(\sigma\in\{\uparrow,\downarrow\}\), define the annihilation operator
\[
\psi_\sigma(f):\mathcal H_1^{\wedge N}\to\mathcal H_1^{\wedge(N-1)}
\]
by
\[
(\psi_\sigma(f)\Psi)^{(N-1)}(x_2,\dots,x_N)
=
\sqrt{N}\int_{\mathbb R^3} d\mathbf r\;
f(\mathbf r)\,
\Psi^{(N)}\big((\mathbf r,\sigma),x_2,\dots,x_N\big),
\]
and define the creation operator \(\psi_\sigma^\dagger(f)\) as the Hilbert-space adjoint of \(\psi_\sigma(f)\).

\begin{mdframed}
  \begin{ques}
  Why is the creation operator the adjoint of the annihilation operator?
\end{ques}
\end{mdframed}

The map \(f\mapsto\psi_\sigma(f)\) is linear in \(f\), and the map
\(f\mapsto\psi_\sigma^\dagger(f)\) is antilinear. Accordingly, there exist operator-valued distributions
\(\psi_\sigma(\mathbf r)\) and \(\psi_\sigma^\dagger(\mathbf r)\) such that
\[
\psi_\sigma(f)
=
\int_{\mathbb R^3} d\mathbf r\; f^{\ast} (\mathbf r)\,\psi_\sigma(\mathbf r),
\qquad
\psi_\sigma^\dagger(f)
=
\int_{\mathbb R^3} d\mathbf r\; f(\mathbf r)\,\psi_\sigma^\dagger(\mathbf r).
\]
These relations express the integral representation of the linear dependence of
\(\psi_\sigma(f)\) and \(\psi_\sigma^\dagger(f)\) on the test function \(f\).

\subsubsection*{F2-3 Particle indistinguishability and slot independence}

Because \(\Psi^{(N)}\) is antisymmetric, the definition of \(\psi_\sigma(f)\) is independent of which coordinate slot is integrated over: integrating over any one slot yields the same \((N-1)\)-particle state up to the sign already fixed by antisymmetry. Thus \(\psi_\sigma(f)\) does not remove a distinguished particle but removes one fermion in the one-particle state specified by \((f,\sigma)\).

\subsubsection*{F2-4 Conditional interpretation of the reduced state}

For any region \(\Omega_{N-1}\subset(\mathbb R^3\times\{\uparrow,\downarrow\})^{N-1}\),
\[
\int_{\Omega_{N-1}}
|(\psi_\sigma(f)\Psi)^{(N-1)}(x_2,\dots,x_N)|^2
\,dx_2\cdots dx_N
\]
is the probability density for finding the remaining \(N-1\) fermions in \(\Omega_{N-1}\),
conditional on one fermion being in the one-particle state \((f,\sigma)\).







\subsubsection*{F2-5 Derivation of the mode expansion of $\psi(f)$}

Assume that the annihilation operator
\[
f \;\mapsto\; \psi(f)
\]
is a linear map from the one-particle Hilbert space $H_1$ to operators on fermionic Fock space.

Let $\{\psi_k\}\subset H_1$ be an orthonormal basis of $H_1$, and define the corresponding mode operators by
\[
c_k \;\equiv\; \psi(\psi_k).
\]

Any $f\in H_1$ admits the expansion
\[
f(x)=\sum_k \psi_k(x)\,\langle \psi_k,f\rangle.
\]

Applying the linear map $\psi(\cdot)$ to this expansion and using linearity,
\[
\psi(f)
=\psi\!\left(\sum_k \langle \psi_k,f\rangle\,\psi_k\right)
=\sum_k \langle \psi_k,f\rangle\,\psi(\psi_k)
=\sum_k \langle \psi_k,f\rangle\,c_k.
\]

The inner product $\langle \psi_k,f\rangle$ can be written in coordinate representation as
\[
\langle \psi_k,f\rangle=\int dx\,\psi_k^{*}(x)\,f(x).
\]

Substituting this expression yields
\[
\psi(f)
=\int dx\, f^{\ast} (x)\,\Big(\sum_k \psi_k(x)\,c_k\Big).
\]



\subsubsection*{F2-6 Canonical anticommutation relations}

The field operators satisfy the canonical anticommutation relations
\[
\{\psi_\sigma(\mathbf r),\psi_{\sigma'}^\dagger(\mathbf r')\}
=
\delta_{\sigma\sigma'}\,\delta(\mathbf r-\mathbf r'),
\qquad
\{\psi_\sigma(\mathbf r),\psi_{\sigma'}(\mathbf r')\}=0,
\qquad
\{\psi_\sigma^\dagger(\mathbf r),\psi_{\sigma'}^\dagger(\mathbf r')\}=0,
\]
to be interpreted after smearing with test functions.

Equivalently, for smeared operators,
\[
\{\psi_\sigma(f),\psi_{\sigma'}^\dagger(g)\}
=
\delta_{\sigma\sigma'}\,\langle f,g\rangle_{\mathcal H_1},
\]
where
\[
\langle f,g\rangle_{\mathcal H_1}
=
\int_{\mathbb R^3} d\mathbf r\; f^*(\mathbf r)g(\mathbf r)
\]
is the inner product on the one-particle Hilbert space.

This relation expresses the consistency condition that creating a fermion in a one-particle state and then testing for its presence returns precisely the overlap of those one-particle states, and nothing else.




\subsubsection*{F2-7 Derivation of the one-body Hamiltonian in field-operator form}

Let $h_0$ be a linear operator on the one-particle Hilbert space $H_1$.
Fix an orthonormal basis $\{\psi_k\}\subset H_1$, and define mode operators
\[
c_k \equiv \psi(\psi_k).
\]
Define the matrix elements of $h_0$ in this basis by
\[
h_{kl} \;\equiv\; \langle \psi_k,\, h_0 \psi_l\rangle_{H_1}.
\]

The corresponding second-quantized one-body Hamiltonian is defined by
\[
H_0 \;\equiv\; \sum_{k,l} h_{kl}\, c_k^\dagger c_l .
\]

Using the mode expansions
\[
\psi(x)=\sum_l \psi_l(x)\,c_l,
\qquad
\psi^\dagger(x)=\sum_k \psi_k^*(x)\,c_k^\dagger,
\]
one computes
\[
\int dx\,\psi^\dagger(x)\,(h_0\psi)(x)
=
\int dx\,
\left(\sum_k \psi_k^*(x)c_k^\dagger\right)
\left(h_0\sum_l \psi_l(x)c_l\right).
\]

Since $h_0$ acts only on the $x$-dependence of $\psi_l(x)$, this becomes
\[
\int dx\,\psi^\dagger(x)\,(h_0\psi)(x)
=
\sum_{k,l}
\left(\int dx\,\psi_k^*(x)\,(h_0\psi_l)(x)\right)
c_k^\dagger c_l
=
\sum_{k,l} h_{kl}\, c_k^\dagger c_l.
\]

Therefore,
\[
H_0
=
\sum_{k,l} \langle \psi_k, h_0 \psi_l\rangle\, c_k^\dagger c_l
=
\int dx\,\psi^\dagger(x)\,h_0\,\psi(x).
\]




\subsubsection*{F2-8. One-body Hamiltonian: \texorpdfstring{$\int dx\,\psi^\dagger(x)\,h_0\,\psi(x)$}{∫ψ† h0 ψ} equals \texorpdfstring{$\sum_{i=1}^N h_0^{(i)}$}{Σ h0(i)} on \texorpdfstring{$\mathcal H_1^{\wedge N}$}{H1^∧N}}

Let \(h_0\) be a one-particle operator on \(\mathcal H_1\) (e.g.\ a differential operator in the coordinate representation).
Define the second-quantized one-body operator on fermionic Fock space by
\[
\widehat H_0 \;\equiv\; \int dx\;\psi^\dagger(x)\,h_0\,\psi(x),
\qquad x=(\mathbf r,\sigma),\quad dx=\sum_{\sigma}\,d^3\mathbf r,
\]
interpreted distributionally via smearing as in F2-2/F2-5.

We show that the restriction of \(\widehat H_0\) to the \(N\)-particle sector \(\mathcal H_1^{\wedge N}\) agrees with the first-quantized operator
\[
H_0^{(N)} \;=\; \sum_{i=1}^N h_0^{(i)},
\]
where \(h_0^{(i)}\) denotes \(h_0\) acting on the \(i\)-th coordinate \(x_i\) of \(\Psi^{(N)}(x_1,\dots,x_N)\).

Let \(\Phi^{(N)},\Psi^{(N)}\in\mathcal H_1^{\wedge N}\). Consider the matrix element
\[
\langle \Phi^{(N)}, \widehat H_0\,\Psi^{(N)}\rangle .
\]
Using adjoints and that \(h_0\) is a one-particle operator acting on the \(x\)-variable,
\[
\langle \Phi^{(N)}, \widehat H_0\,\Psi^{(N)}\rangle
=
\int dx\;
\big\langle
\psi(x)\Phi^{(N)},
\,
h_0\,\psi(x)\Psi^{(N)}
\big\rangle_{\mathcal H_1^{\wedge (N-1)}} .
\]

By the definition of the annihilation operator in coordinate representation (F2-2),
\[
(\psi(x)\Psi^{(N)})^{(N-1)}(x_2,\dots,x_N)
=
\sqrt N\,\Psi^{(N)}(x,x_2,\dots,x_N),
\]
and similarly for \(\Phi^{(N)}\). Substituting this into the previous line and writing the
\((N-1)\)-particle inner product as an integral over \(x_2,\dots,x_N\) gives
\begin{align*}
\langle \Phi^{(N)}, \widehat H_0\,\Psi^{(N)}\rangle
=
N\int dx\,dx_2\cdots dx_N\;
\Phi^{(N)}(x,x_2,\dots,x_N)^{\!*}\;
\big(h_0^{(x)}\Psi^{(N)}\big)(x,x_2,\dots,x_N),
\end{align*}
where \(h_0^{(x)}\) means: apply the one-particle operator \(h_0\) to the \(x\)-dependence of \\
\(\Psi^{(N)}(x,x_2,\dots,x_N)\), holding \(x_2,\dots,x_N\) fixed.
\begin{mdframed}\begin{ques}
  In order for this to be well-defined then, the value of 
\end{ques}

\end{mdframed}


Relabeling \(x\) as \(x_1\),
\[
\langle \Phi^{(N)}, \widehat H_0\,\Psi^{(N)}\rangle
=
N\int dx_1\cdots dx_N\;
\Phi^{(N)}(x_1,\dots,x_N)^{\!*}\;
\big(h_0^{(1)}\Psi^{(N)}\big)(x_1,\dots,x_N).
\]

Because \(\Phi^{(N)}\) and \(\Psi^{(N)}\) are antisymmetric, the integral with \(h_0\) acting on the
first slot has the same value as the integral with \(h_0\) acting on any slot \(i\).
Summing over \(i=1,\dots,N\) therefore produces the factor \(N\), i.e.
\[
N\int dx_1\cdots dx_N\;\Phi^{(N)\!*}\,(h_0^{(1)}\Psi^{(N)})
=
\sum_{i=1}^N \int dx_1\cdots dx_N\;\Phi^{(N)\!*}\,(h_0^{(i)}\Psi^{(N)}).
\]
Hence
\[
\langle \Phi^{(N)}, \widehat H_0\,\Psi^{(N)}\rangle
=
\left\langle
\Phi^{(N)},
\left(\sum_{i=1}^N h_0^{(i)}\right)\Psi^{(N)}
\right\rangle.
\]

Since this holds for arbitrary \(\Phi^{(N)}\), the restriction of \(\widehat H_0\) to \(\mathcal H_1^{\wedge N}\)
equals \(\sum_{i=1}^N h_0^{(i)}\) as an operator on the \(N\)-particle sector.
\subsubsection*{F2-9. Two-body interaction in a mode basis}

Let $\{\varphi_i(x)\}$ be an orthonormal basis of the one-particle Hilbert space
$\mathcal H_1$, where $x=(\mathbf r,\sigma)$, and let
\[
\psi(x)=\sum_i \varphi_i(x)\,c_i,
\qquad
\psi^\dagger(x)=\sum_i \varphi_i^*(x)\,c_i^\dagger
\]
be the corresponding mode expansion of the field operators.

Consider a general two-body interaction kernel
$V(x_1,x_2)$, symmetric under exchange of its arguments.
Define the interaction matrix elements in the mode basis by
\[
V_{ijkl}
\;\equiv\;
\int dx_1\,dx_2\;
\varphi_i^*(x_1)\,\varphi_j^*(x_2)\,
V(x_1,x_2)\,
\varphi_k(x_1)\,\varphi_l(x_2).
\]

These quantities are the matrix elements of the two-body operator
$V$ between antisymmetrized two-particle states constructed from the
one-particle basis $\{\varphi_i\}$.

With this definition, the second-quantized two-body interaction operator
can be written in mode form as
\[
\widehat H_{\mathrm{int}}
=
\frac12
\sum_{i j k l}
V_{ijkl}\,
c_i^\dagger c_j^\dagger c_l c_k .
\]

The factor $\tfrac12$ compensates for the double counting of particle pairs
in the summation over mode indices and corresponds to the restriction
$i\neq j$ in the first-quantized expression
$\tfrac12\sum_{i\neq j}V(x_i,x_j)$.

This expression is purely algebraic once the kernel $V(x_1,x_2)$
and the one-particle basis are fixed; no approximation is involved at this
stage. The equivalence between this mode-space form and the coordinate-space
field-operator form is established by explicit comparison of matrix elements
(see F2-10).

\subsubsection*{F2-10. Equivalence of the two-body interaction in first and second quantization}

We show that the second-quantized operator
\begin{equation}
\widehat H_{\mathrm{int}}
=
\frac12
\int dx_1\,dx_2\;
\psi^\dagger(x_1)\psi^\dagger(x_2)\,
V(x_1,x_2)\,
\psi(x_2)\psi(x_1)
\end{equation}
acts on the \(N\)-particle sector \(\mathcal H_1^{\wedge N}\) as the first-quantized interaction
\[
(H_{\mathrm{int}}^{(N)}\Psi^{(N)})(x_1,\dots,x_N)
=
\frac12\sum_{i\neq j} V(x_i,x_j)\,\Psi^{(N)}(x_1,\dots,x_N).
\]

To avoid ill-defined pointwise products, \(\widehat H_{\mathrm{int}}\) is interpreted distributionally via smearing, as in F2-2 and F2-5.  
We prove equality by comparing matrix elements on the \(N\)-particle sector.

Let \(\Phi^{(N)},\Psi^{(N)}\in\mathcal H_1^{\wedge N}\). Consider
\[
\langle \Phi^{(N)}, \widehat H_{\mathrm{int}} \Psi^{(N)}\rangle .
\]

Using adjoints and the fact that \(V(x_1,x_2)\) is a c-number kernel,
\[
\langle \Phi^{(N)}, \widehat H_{\mathrm{int}} \Psi^{(N)}\rangle
=
\frac12
\int dx_1\,dx_2\;
\big\langle
\psi(x_2)\psi(x_1)\Phi^{(N)},
\,
V(x_1,x_2)\,
\psi(x_2)\psi(x_1)\Psi^{(N)}
\big\rangle_{\mathcal H_1^{\wedge (N-2)}} .
\]

From the definition of the annihilation operator (F2-2),
\[
(\psi(x_2)\psi(x_1)\Psi^{(N)})^{(N-2)}(x_3,\dots,x_N)
=
\sqrt{N(N-1)}\,
\Psi^{(N)}(x_1,x_2,x_3,\dots,x_N),
\]
and similarly for \(\Phi^{(N)}\).

Substituting this into the previous expression and performing the
\((N-2)\)-particle inner product yields
\begin{align*}
&\langle \Phi^{(N)}, \widehat H_{\mathrm{int}} \Psi^{(N)}\rangle
=\\
&\frac12 N(N-1)
\int dx_1\,dx_2\,dx_3\cdots dx_N\;
\Phi^{(N)}(x_1,\dots,x_N)^{\!*}
\,V(x_1,x_2)\,
\Psi^{(N)}(x_1,\dots,x_N).
\end{align*}

Because \(\Phi^{(N)}\) and \(\Psi^{(N)}\) are antisymmetric, the integral with
\(V\) acting on the specific ordered pair \((1,2)\) has the same value as the
integral with \(V\) acting on any ordered pair of distinct slots \((i,j)\).
There are exactly \(N(N-1)\) such ordered pairs. Hence,
\begin{align*}
&N(N-1)
\int dx_1\cdots dx_N\;
\Phi^{(N)\!*}(x_1,\dots,x_N)\,V(x_1,x_2)\,\Psi^{(N)}(x_1,\dots,x_N)
=\\
&\sum_{i\neq j}
\int dx_1\cdots dx_N\;
\Phi^{(N)\!*}(x_1,\dots,x_N)\,V(x_i,x_j)\,\Psi^{(N)}(x_1,\dots,x_N).
\end{align*}

Combining the previous expressions gives
\[
\langle \Phi^{(N)}, \widehat H_{\mathrm{int}} \Psi^{(N)}\rangle
=
\left\langle
\Phi^{(N)},
\left(
\frac12\sum_{i\neq j} V(x_i,x_j)
\right)
\Psi^{(N)}
\right\rangle .
\]

Since this holds for arbitrary \(\Phi^{(N)}\), the operators agree on
\(\mathcal H_1^{\wedge N}\). Thus \(\widehat H_{\mathrm{int}}\) is the
second-quantized representative of the two-body interaction.

\subsubsection*{F2-11. Special case of two body interaction with \(V(x,x') = g_{\alpha,\beta}\delta(\mathbf{r}-\mathbf{r}')\)}
Let \(\mathbf{r_j}\) be the position coordinate of \(x_j\) (which includes spin), and \(g_{\alpha,\beta} = \begin{cases}
  g & \alpha \neq \beta\\
  0 & \alpha = \beta
\end{cases}\).\\ Then
\[
\widehat H_{\mathrm{int}}
=
\frac12
\int dx_1\,dx_2\;
\psi^\dagger(x_1)\psi^\dagger(x_2)\,
g_{\alpha\beta}\delta(\mathbf{r}_1 - \mathbf{r}_2)\,
\psi(x_2)\psi(x_1).
\]
Since the delta function projects this integral onto the subspace where \(\mathbf{r}_1 = \mathbf{r}_2\),
\[
\frac{1}{2}g \sum_{\alpha \neq \beta}\int  \mathrm{d} \mathbf{r}\; \psi_\alpha^{\dagger}(\mathbf{r})\psi_\beta^{\dagger}(\mathbf{r})\psi_\beta(\mathbf{r}) \psi_\alpha(\mathbf{r})
\]



%\subsection*{I. Mathematical and Structural Foundations}

%\begin{enumerate}

%\item \textbf{Tensor product structure.}  
%The many-body Hilbert space is constructed as a graded tensor product of one-particle Hilbert spaces, with fermionic antisymmetrization imposed.
%\begin{ques}
  %What is a tensor product of spaces? What does it mean that fermionic antisymmetrization is imposed? I understand that "swapping particles" must result in negation of the wave function, but what is the mathematical definition of "swapping particles" - I recall that with two particles, it is realized by negating the coordinate system, but am not sure how this extends to the tensor product and many particles. \end{ques}
  
  %\begin{ans}
    %I re-consulted my linear algebra textbook for the formal definition, and physically, the tensor product of states \(\phi\text{ and } \tau\) is the state of two particles where particle 1 is in state \(\phi\) and particle 2 is in state \(\tau\). \\
  %This is a tensor product of \(\mathcal{H}\) with itself, so swapping particles is simply permuting factors in the tensor product. Hence "anticommutativity" of the \(\otimes\) operator. \end{ans}
    %\begin{ques}
    %\(\psi(r_1,r_2)\) is shorthand for \(\psi_1(r) \otimes \psi_2(r)\)? But is then \(\psi(r_1,r_2)\) no longer a map from \(R^3\) to \(\C\), but rather from \(\R^6\) (not considering the spinor)? Is the "ansatz" wave function \(\tilde{\psi}\) a map from some huge configuration space, or just \(R^3\)? I guess if \(\psi\) represents a state for which the particle number is uncertain, then it would not even be certain what \(\psi \) is a map from. Perhaps the interpretation of wave functions as functions breaks down under the application of the tensor product, and meaning is only extracted through field operators.
  %\end{ques}
  %\begin{ques}
    %What is a graded tensor product? From what I read in LADR, it seems that given two vector spaces, there is just one way to take the tensor product: \(v \otimes w\) is the functional so that \((v \otimes w)(\phi, \tau) = \phi(v)\tau(w)\), and in this case, it is not clear that the tensor product is anticommuting. 
  %\end{ques}
 
%\item \textbf{Operator-valued distributions.}  
%Field operators $\psi(\mathbf r)$ and $\psi^\dagger(\mathbf r)$ are operator-valued distributions. Products at coincident points are only meaningful under integrals or suitable regularization.
%\begin{ques}
%Are coincident points just identical points?
%\end{ques}
%\end{enumerate}

%\subsection*{II. Second Quantization: Formal Axioms}

%\begin{enumerate}
%\setcounter{enumi}{3}

%\item \textbf{Fermionic Fock space.}  
%The fermionic Fock space is constructed from the one-particle Hilbert space. The vacuum state $\lvert 0\rangle$ is annihilated by all annihilation operators.
%\begin{ques}
    %How is it constructed?
%\end{ques}

%\item \textbf{Canonical anticommutation relations (CAR).}
%\[
%\{\psi_\alpha(\mathbf r),\psi_\beta^\dagger(\mathbf r')\}
%= \delta_{\alpha\beta}\delta(\mathbf r-\mathbf r'), \qquad
%\{\psi,\psi\}=\{\psi^\dagger,\psi^\dagger\}=0.
%\]
%\begin{ques}
  %These are related to the fermionic particle swapping negation rules right? How?
%\end{ques}

%\item \textbf{Operator correspondence.}  
%One-body operators correspond to bilinear forms in field operators; two-body operators correspond to quartic forms. Normal ordering is defined relative to the vacuum.

%\begin{ques}
  %What is normal ordering?
%\end{ques}
%\item \textbf{Density and number operators.}
%\[
%\hat n(\mathbf r)=\sum_\alpha \psi_\alpha^\dagger(\mathbf r)\psi_\alpha(\mathbf r).
%\]
%This is taken as an operator identity. 
%\begin{ques}
  %\(\int_V \hat{n}(\mathbf{r}) \mathrm{d}\mathbf{r}\) is the expectation value for the number of particles in volume \(V\)? 
%\end{ques}
%\end{enumerate}

%\subsection*{III. Microscopic Hamiltonian Assumptions}

%\begin{enumerate}
%\setcounter{enumi}{7}

%\item \textbf{Electronic Hamiltonian.}  
%The Hamiltonian consists of kinetic energy, an external potential, and an instantaneous two-body interaction. Spin indices are included explicitly; spin--orbit coupling is absent unless stated.

%\item \textbf{Attractive interaction channel.}  
%Only the interaction channel relevant for pairing is retained. This is a controlled truncation, not an identity.
%\end{enumerate}

%\subsection*{IV. Statistical Mechanics Layer}

%\begin{enumerate}
%\setcounter{enumi}{9}

%\item \textbf{Grand canonical ensemble.}  
%The system is described in the grand canonical ensemble with chemical potential $\mu$ included in the Hamiltonian.

%\item \textbf{Thermodynamic limit.}  
%The limit $V\to\infty$ is taken prior to symmetry considerations, allowing inequivalent representations of the CAR.
%\begin{ques}
  %What is \(V\), and what are inequivalent representations of the CAR?
%\end{ques}

%\item \textbf{Expectation values.}  
%Expectation values $\langle \cdot \rangle$ denote ensemble averages in a chosen equilibrium state and yield c-numbers.
%\begin{ques}
  %What is a c-number?
%\end{ques}
%\end{enumerate}

%\subsection*{V. Mean-Field / Saddle-Point Commitment}

%\begin{enumerate}
%\setcounter{enumi}{12}

%\item \textbf{Mean-field factorization.}  
%Quartic operator products may be approximated by bilinear terms plus c-number fields via a saddle-point approximation. 
%\begin{ques}
  %What is a saddle=point approximation, and why is this an axiom? Doesn't Svidzinsky go through this approximation somewhat carefully?
%\end{ques}

%\item \textbf{Order parameter definition.}
%\[
%\Delta(\mathbf r,\mathbf r')
%\equiv
%- V(\mathbf r-\mathbf r')
%\langle \psi_\downarrow(\mathbf r')\psi_\uparrow(\mathbf r)\rangle.
%\]
%This is a definition.
%\begin{ques}
%We're working over momentum states, not position states, so should these \(\mathbf{r}\) be replaced by \(\mathbf{k}\). If not, what does \(V(\mathbf{r} - \mathbf{r}')\) mean physically, considering particles are not localized?
%\end{ques}

%\item \textbf{Self-consistency condition.}  
%The mean field must equal the expectation value computed in the resulting quadratic Hamiltonian. 
%\begin{ques}
  %Define the mean field, expectation value of what?
%\end{ques}
%\end{enumerate}

%\subsection*{VI. Symmetry-Breaking Commitments}

%\begin{enumerate}
%\setcounter{enumi}{15}

%\item \textbf{Broken $U(1)$ symmetry.}  
%Particle-number conservation is not enforced at the level of the effective Hamiltonian.
%\begin{ques}
  %Does this mean that the Hamiltonian can change the number of particles in the system? What would this even mean physically?
%\end{ques}

%\item \textbf{Anomalous averages.}  
%Expectation values of the form $\langle \psi\psi\rangle$ are permitted and nonzero in the chosen symmetry-broken sector. 
%\begin{ques}
  %What is a symmetry-broken sector? I assume \(\psi \) are the field operators. Why might these not be permitted? Permitted is a rather heuristic term.
%\end{ques}

%\item \textbf{Gauge fixing.}  
%A definite global phase of the order parameter is implicitly fixed. 
%\begin{ques}
  %What is a global phase, and what is an order parameter?
%\end{ques}
%\end{enumerate}

%\subsection*{VII. Quadratic Fermionic Hamiltonians}

%\begin{enumerate}
%\setcounter{enumi}{18}

%\item \textbf{Quadratic Hamiltonians.}  
%Any fermionic Hamiltonian quadratic in field operators admits exact diagonalization.

%\item \textbf{Bogoliubov transformation.}  
%Diagonalization is achieved via a linear canonical transformation preserving the CAR.

%\item \textbf{Nambu formalism.}  
%Doubling of degrees of freedom is a bookkeeping device and introduces no new physical assumptions.
%\end{enumerate}

%\subsection{Second Quantization}
%\begin{enumerate}
%\item \textbf{Fermionic modes and CAR.}
%For any single-particle label $a$ (e.g. $(\mathbf r,\sigma)$ or $(\mathbf k,\sigma)$),
%there exist operators $c_a, c_a^\dagger$ obeying
%$\{c_a,c_b^\dagger\}=\delta_{ab}$ and $\{c_a,c_b\}=\{c_a^\dagger,c_b^\dagger\}=0$.

%\item \textbf{Field operators as a basis change.}
%The real-space fields are linear combinations of mode operators:
%$\psi_\sigma(\mathbf r)=\sum_a \phi_a(\mathbf r,\sigma)\,c_a$,
%$\psi_\sigma^\dagger(\mathbf r)=\sum_a \phi_a^*(\mathbf r,\sigma)\,c_a^\dagger$,
%with $\{\psi_\sigma(\mathbf r),\psi_{\sigma'}^\dagger(\mathbf r')\}
%=\delta_{\sigma\sigma'}\delta(\mathbf r-\mathbf r')$.

%\item \textbf{Many-body Hamiltonians in second quantization.}
%One-body terms are quadratic in fields/modes; two-body terms are quartic.
%In particular, a contact interaction is
%\[
  %H_{\mathrm{int}}=g\int d\mathbf r\;
  %\psi_\uparrow^\dagger(\mathbf r)\psi_\downarrow^\dagger(\mathbf r)
  %\psi_\downarrow(\mathbf r)\psi_\uparrow(\mathbf r)
%\]

%\item \textbf{Grand-canonical ensemble as trace over Fock space.}
%Thermodynamics is encoded by
%$Z=\mathrm{Tr}\,e^{-\beta(H-\mu\hat N)}$,
%where $\hat N$ is the number operator.

%\item \textbf{Interaction picture in imaginary time.}
%Given $\mathcal H_0=H_0-\mu\hat N$, define
%$O(\tau)=e^{\tau\mathcal H_0}Oe^{-\tau\mathcal H_0}$,
%and the time-ordering operator $T_\tau$ orders operators by $\tau$.

%\item \textbf{Time-ordered exponential identity.}
%For an interaction $H_{\mathrm{int}}(\tau)$ in the interaction picture,
%$Z=\mathrm{Tr}\left\{e^{-\beta\mathcal H_0}\,
%T_\tau\exp\left(-\int_0^\beta H_{\mathrm{int}}(\tau)\,d\tau\right)\right\}$.

%\item \textbf{Gaussian (Wick) property for quadratic fermions.}
%Thermal averages with respect to a quadratic Hamiltonian obey Wick’s theorem:
%time-ordered products reduce to sums of products of two-point contractions.

%\item \textbf{Hubbard--Stratonovich (HS) identity (operator form).}
%For an operator bilinear $L(\mathbf r,\tau)$,
%\[
%T_\tau e^{\int L^\dagger L}
%\propto
%\int D\zeta D\zeta^*\,
%e^{-\int |\zeta|^2}\,
%T_\tau e^{\int(L^\dagger\zeta+\zeta^*L)} .
%\]
%(Precise constants are fixed by normalization.)

%\item \textbf{Auxiliary pair field scaling.}
%If $L\propto \sqrt{|g|}$, then scaling $\zeta=|g|^{-1/2}\Delta$
%rewrites $Z$ as a functional integral over $\Delta(\mathbf r,\tau)$.

%\item \textbf{Saddle-point/mean-field principle.}
%When the functional integral is dominated by stationary configurations,
%$\Delta(\mathbf r,\tau)$ may be replaced by a stationary (often $\tau$-independent)
%field $\Delta(\mathbf r)$ determined by the stationarity condition.

%\item \textbf{Quadratic mean-field Hamiltonian.}
%At fixed $\Delta(\mathbf r)$ the fermionic problem is quadratic and can be written
%as
%\[
%H_{\mathrm{MF}}=
%\sum_{\sigma}\int \psi_\sigma^\dagger h_0 \psi_\sigma\,d\mathbf r
%+\int\left(\Delta\,\psi_\uparrow^\dagger\psi_\downarrow^\dagger
%+\Delta^*\,\psi_\downarrow\psi_\uparrow\right)\,d\mathbf r
%+\text{(c-number)}.
%\]

%\item \textbf{Bogoliubov--Valatin (canonical) transformation.}
%There exist quasiparticle operators $\gamma_n,\gamma_n^\dagger$
%(linear in $\psi,\psi^\dagger$) such that:
%(i) they satisfy CAR, and
%(ii) they diagonalize $H_{\mathrm{MF}}$.
%The CAR constraint implies the usual orthonormality/completeness relations
%for the coefficients $(u_n(\mathbf r),v_n(\mathbf r))$.

%\item \textbf{BdG eigenproblem from diagonalization.}
%Requiring that $H_{\mathrm{MF}}$ be diagonal in $\gamma_n^\dagger\gamma_n$
%yields a coupled linear eigenvalue problem for $(u_n,v_n)$
%(the Bogoliubov--de Gennes equations), plus the self-consistency condition
%$\Delta(\mathbf r)=-g\langle \psi_\downarrow(\mathbf r)\psi_\uparrow(\mathbf r)\rangle$.
%\end{enumerate}



\section{January 23, 2026}
\noindent\textbf{Tasks} 
\begin{itemize}
  \item Finish studying foundational field operator and second quantization material
  \item Begin studying first part of Svidzinsky derivation of approximation for interaction.
\end{itemize}




\end{document}
